Ultrafast Electron Microscopes: Design Criteria, Electron Sources, and Column Modeling

### 2013-04-05 inc/model_results/introduction.tex
merge more comments from andreas, now on to model_results chapter

In Chapters \ref{chap:previous_models} and \ref{chap:extension}, I have introduced the extended Analytic Gaussian model.
In the following chapter I use this model to make general observations about pulse propagation for Ultrafast Electron Microscopy.

These examples will each use one of two initialization schemes.
The first represents an idealized pulse which is not generated by photocathode and then accelerated, but rather springs into existence at its initial position with its velocity already established.
This type of pulse is useful to demonstrate general pulse dynamics or the ideal influence of electron-optical elements independent of the pulse's creation conditions.
Simulations using pulses of this type will provide an initial velocity $v_{\smallzero}$ (or equivalent acceleration voltage $V_{DC}$), an excess photoemission energy $\Delta E$, and HW1/eM dimensions width $w$ and duration $\tau$ or length $ w_z = v_{\smallzero} \tau$.
The width and length will be used directly to calculate initial spatial variances $\sigma_{\alpha}$.
Both momentum chirp terms, $\gamma_{\alpha}$, will be initially zero, but since pulses must still be subject to Liouville's Theorem (discussed in Section \ref{sec:liouville}), \ref{eq:summary} will still be used to calculate meaningful local momentum widths $\eta_{\alpha}$.
This idealization only applies to the initialization, therefore these pulses will still be subject to intrinsic pulse broadening and internal space-charge repulsion which might cause pulse expansion before or during the demonstration, especially for high charge densities.

>>>>

In Chapters \ref{chap:previous_models} and \ref{chap:extension}, I have introduced the extended Analytic Gaussian model.
In this chapter, I use this model to make general observations about pulse propagation for Ultrafast Electron Microscopy.

The examples presented will each use one of two initialization schemes.
The first represents an idealized pulse which is not generated by photocathode and then accelerated, but rather springs into existence at its initial position with its velocity already established.
This type of pulse is useful to demonstrate general pulse dynamics or the ideal influence of electron-optical elements independent of the pulse's creation conditions.
Simulations using pulses of this type will provide an initial velocity $v_{\smallzero}$ (or equivalent acceleration voltage $V_{DC}$), an excess photoemission energy $\Delta E$, and HW1/eM dimensions; width $w$ and duration $\tau$ or length $ w_z = v_{\smallzero} \tau$.
The width and length will be used directly to calculate initial spatial variances $\sigma_{\alpha}$.
Both momentum chirp terms, $\gamma_{\alpha}$, will be initially zero, but since pulses must still be subject to Liouville's Theorem (discussed in Section \ref{sec:liouville}), \ref{eq:summary} will still be used to calculate meaningful local momentum widths $\eta_{\alpha}$.
This idealization only applies to the initialization, therefore these pulses will still be subject to intrinsic pulse broadening and internal space-charge repulsion which might cause pulse expansion before or during the demonstration, especially for high charge densities.

### 2013-04-05 inc/model_results/shapes/shapes.tex
merge more comments from andreas, now on to model_results chapter

The initial `space-time' eccentricity (defined in \ref{eq:define_xi}),
\begin{equation}
  \xi ( 0 ) = \sqrt{ \frac{ \sigma_{z}(0) }{ \sigma_{\smallT}(0) } } = \frac{ w_z(0) }{ w(0) } = \frac{ v_{\smallzero} \tau(0) }{ w(0) } \,\text{,}
\end{equation}
of the photo-generated electron pulse plays a major role in the bunch propagation dynamics, particularly when space-charge effects dominate.
\ref{fig:compare_shape} illustrates the typical dynamics of initially spherical ($ \xi(0) = 0 $), oblate or `disk-like' ($ \xi(0) < 1 $), and prolate or `cigar-like' ($ \xi(0) > 1 $) pulses.
%TODO rewrite using w?
This simulation uses the idealized pulse conditions described above, where the initial bunch dimensions are given by the ellipticities $ \xi(0) = \{ 1/3 , 1 , 3 \} $ and in each case the product of the initial HW1/eM bunch widths $w^2 w_z = $ 1mm$^3$.
Pulses used in \ref{fig:compare_shape}\subref{fig:compare_shape_N1} have just one electron per pulse (i.e. $ N = 10^{0} $), while \ref{fig:compare_shape}\subref{fig:compare_shape_N1e6} have a much higher $ N = 10^{6} $ electrons in the bunch.
Since the initial pulse volume is consistent between all runs, initial peak pulse charge density is six orders of magnitude greater in \ref{fig:compare_shape}\subref{fig:compare_shape_N1e6} than in \ref{fig:compare_shape}\subref{fig:compare_shape_N1}.
The pulse velocity is $ v_{\smallzero} \approx c/3 $, i.e. a DC photogun with $V = 30\text{kV} $.

@@

For an electron pulse with a higher initial space-charge density, intra-pulse Coulomb effects may dominate the propagation dynamics.
\ref{fig:compare_shape}\subref{fig:compare_shape_N1e6}, shows the predicted time evolution dynamics of electron pulses with the same initial size and photoemission conditions as \ref{fig:compare_shape}\subref{fig:compare_shape_N1}, but with $ N = 10^{6} $ electrons --- the number required for single-shot diffraction pattern measurements in a DTEM.\cite{berger_dc_2009,armstrong_practical_2007}
The significant influence of space-charge effects for the higher initial peak pulse charge density is clearly evident.
In the spherical case, space-charge induced bunch broadening begins to overwhelm the initial difference between $\eta_{\smallT}$ and $\eta_{z}$, producing a more uniform three-dimensional `Coulomb explosion' of the electron pulse with the overall bunch broadening rate much greater than that for $ N = 10^{0} $ (\ref{fig:compare_shape}\subref{fig:compare_shape_N1}).
For $ \xi (0) > 1 $ (or $ \xi (0) < 1 $), the smaller initial transverse (longitudinal) bunch dimension experiences the greatest relative expansion in free-space propagation. This is to be expected, because the initial $\eta_{\alpha}$ are non-zero (but comparable in magnitude) while Coulomb effects will produce differential pulse broadening in the dimension with the largest internal space-charge field (i.e., the dimension with the largest electrostatic potential gradient).
% the following sentance is no longer true, still perhaps should include something like it
%In fact, both the $ N = 10^{6} $ oblate and prolate pulse examples shown in \ref{fig:compare_shape}\subref{fig:compare_shape_N1e6} transition through a spherical bunch shape ($ \xi \approx 1 $) after a time of flight of $\sim$0.5ns, at which point the peak pulse charge density has already been reduced by a factor of $\sim$100.
The initial impulse induced by the strong space-charge field then provides for continued linear expansion (in time), resulting in a linear pulse chirp in the dimension associated with the smallest initial HW1/eM bunch width\cite{michalik_analytic_2006,berger_dc_2009} $ \sqrt{ 2 \sigma_{\alpha} (0) } $.

>>>>

The initial `space-time' eccentricity (defined in \ref{eq:define_xi}),
\begin{equation}
  \xi ( 0 ) = \sqrt{ \frac{ \sigma_{z}(0) }{ \sigma_{\smallT}(0) } } = \frac{ w_z(0) }{ w(0) } = \frac{ v_{\smallzero} \tau(0) }{ w(0) } \,\text{,}
\end{equation}
of the photo-generated electron pulse plays a major role in the bunch propagation dynamics, particularly when space-charge effects dominate.
\ref{fig:compare_shape} illustrates the typical dynamics of initially spherical ($ \xi(0) = 1 $), oblate or `disk-like' ($ \xi(0) < 1 $), and prolate or `cigar-like' ($ \xi(0) > 1 $) pulses.
%TODO rewrite using w?
This simulation uses the idealized pulse conditions described above, where the initial bunch dimensions are given by the ellipticities $ \xi(0) = \{ 1/3 , 1 , 3 \} $ and in each case the product of the initial HW1/eM bunch widths $w^2 w_z = $ 1mm$^3$.
Pulses used in \ref{fig:compare_shape}\subref{fig:compare_shape_N1} have just one electron per pulse (i.e. $ N = 10^{0} $), while \ref{fig:compare_shape}\subref{fig:compare_shape_N1e6} have a much higher $ N = 10^{6} $ electrons in the bunch.
Since the initial pulse volume is consistent between all runs, initial peak pulse charge density is six orders of magnitude greater in \ref{fig:compare_shape}\subref{fig:compare_shape_N1e6} than in \ref{fig:compare_shape}\subref{fig:compare_shape_N1}.
The pulse velocity is $ v_{\smallzero} \approx c/3 $, which is equivalent to a DC photogun with $V_{DC} = 30\text{kV} $.

@@

For an electron pulse with a higher initial space-charge density, intra-pulse Coulomb effects may dominate the propagation dynamics.
\ref{fig:compare_shape}\subref{fig:compare_shape_N1e6}, shows the predicted time evolution dynamics of electron pulses with the same initial size and photoemission conditions as \ref{fig:compare_shape}\subref{fig:compare_shape_N1}, but with $ N = 10^{6} $ electrons --- the number required for single-shot diffraction pattern measurements in a DTEM.\cite{berger_dc_2009,armstrong_practical_2007}
The significant influence of space-charge effects for the higher initial peak pulse charge density is clearly evident.
In the spherical case, space-charge induced bunch broadening begins to overwhelm the initial difference between $\eta_{\smallT}$ and $\eta_{z}$, producing a more uniform three-dimensional `Coulomb explosion' of the electron pulse with the overall bunch broadening rate much greater than that for $ N = 10^{0} $ (\ref{fig:compare_shape}\subref{fig:compare_shape_N1}).
For $ \xi (0) > 1 $ (or $ \xi (0) < 1 $), the smaller initial transverse (longitudinal) bunch dimension experiences the greatest relative expansion in free-space propagation. This is to be expected, because the initial $\eta_{\alpha}$ are non-zero (but comparable in magnitude) while Coulomb effects will produce the greatest differential pulse broadening in the dimension with the largest internal space-charge field (i.e., the dimension with the largest electrostatic potential gradient).
% the following sentance is no longer true, still perhaps should include something like it
%In fact, both the $ N = 10^{6} $ oblate and prolate pulse examples shown in \ref{fig:compare_shape}\subref{fig:compare_shape_N1e6} transition through a spherical bunch shape ($ \xi \approx 1 $) after a time of flight of $\sim$0.5ns, at which point the peak pulse charge density has already been reduced by a factor of $\sim$100.
The initial impulse induced by the strong space-charge field then provides for continued linear expansion (in time), resulting in a linear pulse chirp in the dimension associated with the smallest initial HW1/eM bunch width\cite{michalik_analytic_2006,berger_dc_2009} $ \sqrt{ 2 \sigma_{\alpha} (0) } $.

### 2013-04-05 inc/model_results/spacecharge/spacecharge.tex
merge more comments from andreas, now on to model_results chapter

As seen in Section \ref{sec:initial_shapes}, at low pulse charge densities (low number of electrons/pulse), the propagation dynamics should be largely determined by the effects associated with the intrinsic geometric pulse broadening; for most photoemission processes this is related to the excess photoemission energy $\Delta E$.
On the other hand, space-charge effects are expected to dominate at large pulse charge densities.
A full analysis of how high the charge density may be before deleterious space-charge become unmanagable would be a sizable task.
While much discussion could be had on the degree of distortion allowable as well as considerations of the initial shape of the pulse and length and arrangement of the column and its constituent elements, a certain amount of intuition can be gained simply by considering a progressive series of examples.

>>>>

As seen in Section \ref{sec:initial_shapes}, at low pulse charge densities, the propagation dynamics should be largely determined by the effects associated with the intrinsic geometric pulse broadening; for most photoemission processes this is related to the excess photoemission energy $\Delta E$.
On the other hand, space-charge effects are expected to dominate at large pulse charge densities.
A full analysis of how high the charge density may be before deleterious space-charge become unmanagable would be a sizable task.
While much discussion could be had on the degree of distortion allowable as well as considerations of the initial shape of the pulse and length and arrangement of the column and its constituent elements, a certain amount of intuition can be gained simply by considering a progressive series of examples.
